Manuscript_Github_Pchabaudi_pir

### 2022-07-27 chapter4_results_chabaudi24h/Thesis_Github_Pchabaudi_pirAnalysis.Rmd
Start moving phase/GO enrichment code to its own Rmd file

```{r gene_sync}

#Which genes are out of sync between rMT and SBP?

#Melt the z-score dataframe into long form, then slice out the maximum z-score value.
# Filter out any gene which has multiple max values for simplicity.
# Summarise this with the numbers that each stage (the max stage) corresponds to.

z_max <- lapply(chab_z.list, function(x) {
  x$zscores %>% 
  reshape2::melt(id.vars = 'Geneid', 
       variable.name = 'sample', 
       value.name = 'zscore') %>% 
  group_by(Geneid) %>% 
  slice_max(order_by = zscore, n = 1) %>% 
  filter(n() == 1) %>% 
  summarise(sample_max = times_stages_table$num[
    str_detect(sample, 
               times_stages_table$time)
    ])
})

#Bind the two together with inner_join so that it only includes genes in both.
# Then calculate the peak difference by subtracting the difference between the sample numbers.
# Since this is a cycle, no point of the cycle is theoretically 
#  more than 4 time points away from the next one. 
#  e.g. 23h and 02h (8 and 1) are only 1 time point apart, not seven.
#  Hence we subtract values above four from eight
z_max_diff <- inner_join(z_max[[1]], z_max[[2]], by = 'Geneid') %>% 
  mutate(diff = ifelse(abs(sample_max.x - sample_max.y) > 4,
                       8 - abs(sample_max.x - sample_max.y),
                       abs(sample_max.x - sample_max.y)))

#Probably unnecessarily complicated code to make a barplot of how many genes 
# are peaking at different timepoints between rMT and SBP
barplot(group_by(z_max_diff, diff) %>% summarise(total = n()) %>% .$total,
        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'))

#And for the pir genes
barplot(group_by(z_max_diff, diff) %>% 
          filter(Geneid %in% cir_id) %>% 
          summarise(total = n()) %>% 
          .$total,
        names.arg = paste0(sort(unique(z_max_diff$diff))*3, 'h'))

```{r phase_go_enrich}
#Get the gene list for each group of genes in and out of phase.
gene.list <- chabaudi_gene_info$Geneid
z_max_diff.list <- lapply(seq(0,4), 
                          function(diff_num) {
                            diff.list <- filter(z_max_diff, diff == diff_num) %>% 
                              .$Geneid
                            #Get 1 for on the list and 0 for not.
                            gene_diff.list <- as.integer(gene.list %in% diff.list)
                            names(gene_diff.list) <- gene.list
                            return(gene_diff.list)
                          })
names(z_max_diff.list) <- seq(0,4)

#Read in the GO-Annotation file using GAFReader from FoldGO.
# getAnnotation to obtain a list of the GO terms and associated genes.
# inverseList (from topGO) to turn this into list of geneID with associated GO terms.
gaf <- GAFReader("../standard_files/PlasmoDB-48_Pchabaudichabaudi_GO.gaf")
go_annotation <- inverseList(getAnnotation(gaf))

# How many genes for each group do not have any GO annotations?

lapply(z_max_diff.list,
       function(gene_diff.list){
         total_noGO <- sum(!names(gene_diff.list[gene_diff.list == 1]) %in% 
                             names(go_annotation))
         prop_noGO <- total_noGO/sum(gene_diff.list == 1)
         return(paste0(total_noGO, ' genes without annotation = ',
                       round(prop_noGO*100, digits = 2), '%'))
       })

```{r}
#Load for the showSigOfNodes network function.
library('Rgraphviz')

#Make a topGOdata object 
GOdata <- new("topGOdata", 
              ontology = "BP", 
              geneSel = function(x){x == 1},
              allGenes = z_max_diff.list$`1`,
              nodeSize = 5,
              annot = annFUN.gene2GO, 
              gene2GO = go_annotation)
resultFis <- runTest(GOdata, 
                     algorithm = "classic", 
                     statistic = "fisher")
plot.new()
hist(score(resultFis), 50, xlab = "p-values")
dev.off()
showSigOfNodes(GOdata,
               score(resultFis),
               firstSigNodes = 5,
               useInfo = 'all')
goRes <- GenTable(GOdata, 
         classic = resultFis, 
         orderBy = "classic", 
         ranksOf = "classic",
         topNodes = 25) %>% 
  filter(classic < 0.05)

#Get first (most significant) enriched GO id 
goID <- goRes[4, "GO.ID"]
#Show distribution of GO term genes compared to null distribution (all other genes)
print(showGroupDensity(GOdata, goID, ranks = TRUE))

#Unlist the lapply below which goes through each element of the list and 
# returns TRUE is the go_id is listed for that gene.
#This makes a TRUE/FALSE index for the list/names of the list which can get 
# the genes associated with that GO id.
goID_genes <- names(go_annotation)[unlist(
  lapply(go_annotation, 
         function(go_ids){
           any(goID %in% go_ids)
         })
)]

draw_heatmap_zscore(zscore_list = chab_z.list,
                    filter_genes_vector = names(z_max_diff.list$`1`[z_max_diff.list$`2` == 1]),
                    gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                    threshold_tpm = 1,
                    use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = TRUE)
```

```{r}

lapply(seq(0,4), 
       function(phase){
         GOdata <- new("topGOdata", 
                       ontology = "BP", 
                       geneSel = function(x){x == 1},
                       allGenes = getElement(z_max_diff.list, paste(phase)),
                       nodeSize = 5,
                       annot = annFUN.gene2GO, 
                       gene2GO = go_annotation)
         resultFis <- runTest(GOdata, 
                              algorithm = "classic", 
                              statistic = "fisher")
         draw_heatmap_zscore(zscore_list = chab_z.list,
                             filter_genes_vector = names(
                               getElement(z_max_diff.list, paste(phase))[
                                 getElement(z_max_diff.list, paste(phase)) == 1
                               ]
                             ),
                             gene_ordering_vector = phaseogram_time_order$rMT$zscore_order$Geneid,
                             threshold_tpm = 1,
                             use_times = TRUE,
                    use_order_vector = FALSE,
                    show_row_names = FALSE,
                    tpm_mat = TRUE)
         GenTable(GOdata, 
                  classic = resultFis, 
                  orderBy = "classic", 
                  ranksOf = "classic",
                  topNodes = 25) %>% 
           filter(classic < 0.05)
       })

@@

Fore node visualisation:
"Rectangles indicate the 5 most signicant terms. Rectangle color represents the relative signicance,
ranging from dark red (most signicant) to bright yellow (least signicant). For each node, some basic information
is displayed"

>>>>

@@

For node visualisation:
"Rectangles indicate the 5 most significant terms. Rectangle color represents the relative significance,
ranging from dark red (most significant) to bright yellow (least significant). For each node, some basic information
is displayed"
